Efficient Compilation of High Level Python Numerical Programs with Pythran

### 2014-08-27 doc/papers/cgo2014/paper.tex
Import some description from Scipy article

@@

@@

\subsection{Numba}
\cite{numba}

\subsection{Parakeet}
\cite{parakeet2012}

@@

@@

>>>>

The success of Python and the development of its large ecosystem triggered
several attempts to improve its execution speed. The most significant basic 
block for high performance numerical simulations is certainly Numpy, a 
numerical library to manipulate dense arrays. Numerous routines are exposed in 
Python and implemented in C, offering close-to-native performance in some cases.
% FIXME: Numpy is part of this section but somehow covered earlier???
% 
However, this library-like approach has inherent limitations that were exposed 
in the previous section. To address these limitations, several compiler-based 
approaches were proposed.

@@

% Not sure what to add here because of the previous section.

@@

% 
Cython~\cite{cython2010} is an hybrid Python/C dialect. It extends the Python
syntax with typing information, calls to native functions from third party
libraries, and a limited set of parallelism constructs, such as the possibility
to define parallel loops, but no task parallelism. When possible, it unboxes
Python variables to improve performance. Without type annotations, the
performance improvement is not terrific, but given enough type annotation,
Cython can generates code that runs as fast as its C equivalent, while
maintaining a syntax close to Python.

\subsection{Numba and Parakeet}

The numba~\footnote{\emph{cf}.\ \texttt{https://github.com/numba/numba}}~\cite{numba}
compiler uses additional type information to generate sequential LLVM bytecode.
Parakeet~\cite{parakeet2012} follows an approach similar to numba, that is
Python to LLVM bytecode translation, but limits its scope to the numpy package,
only supporting a small subset of the Python language. Additionally, it supports
implicit parallelism using an implicit mapping between numpy functions and a set
of parallel primitives including \texttt{map}s \texttt{scan}s and
\texttt{reduce}s. These two approaches use Just-In-Time(JIT) compilation and do
not suffer from backward-incompatibility issues. When meeting an unsupported
construct, numba falls back to Python C-API calls, while parakeet raises an
exception.

@@

Copperhead~\cite{copperhead2011} is a functional, data parallel language
embedded in Python. It uses $n$-uplets, NumPy arrays and lists as its core data
structure and prohibits usage of many control-flow operators such as loops,
enforcing the use of the \texttt{map}, \texttt{filter} or \texttt{reduce}
intrinsics to exhibit parallelism. But it can be efficiently compiled to either
CUDA or C++ with calls to the Thrust\footnote{\emph{cf}.
\texttt{http://thrust.github.com/}} library. Python decorators are used to
identify hot-spots that are JIT-compiled to native code.

@@

Tools such as PyPy~\cite{pypy2009}, a Python interpreter with a tracing JIT, or
Shed~Skin~\cite{shedskin2006}, a Python to C++ compiler are also viable ways to
enhance Python performance. However Shed~Skin does not provide support for
fine-grained parallelism beyond what the standard library proposes. PyPy is
heading toward STM for parallelism support.
